%!TEX root = ../thesis.tex
%*******************************************************************************
%*********************************** First Chapter *****************************
%*******************************************************************************

\chapter{Introduction}  %Title of the First Chapter

\ifpdf
    \graphicspath{{Chapter1/Figs/Raster/}{Chapter1/Figs/PDF/}{Chapter1/Figs/}}
\else
    \graphicspath{{Chapter1/Figs/Vector/}{Chapter1/Figs/}}
\fi

At a policy level, the Tallinn Declaration, signed on 6th October 20172
, confirms the commitment to the vision
laid out in the EU eGovernment Action Plan 2016-20203 and in the European Interoperability Framework
(EIF)4
. In the next five years (2018-2022), steps will be taken towards the declared principles in EU public
administrations, namely: “digital-by-default, inclusiveness and accessibility”, “once-only”, “trustworthiness and
security”, “openness and transparency”, and “interoperability by default”, as well as national interoperability
frameworks based on the European Interoperability Framework (EIF).
In the Declaration the “user-centricity principles for design and delivery of digital public services” is key.
When interacting with public administrations and using digital public services, citizens and businesses
should have: digital interaction, accessibility, security, availability and usability, reduction of the
administrative burden, digital delivery of public services, citizen engagement, incentives for digital service
use, protection of personal data and privacy, redress and complaint mechanisms. Whilst not specifically
covered within the scope of this study, AI has the potential to improve public services and contribute to the
objectives set out in the Tallinn Declaration. For example, the Commission will look into AI's potential to
analyse large amounts of data and help check how single market rules are applied5
.
Moreover, the digital transformation of society, business and government is raising issues for a range of
policy matters across the European Union. As e-government has been in place for the last 20 years, it is
timely to explore the interplay between technology and government activities from the perspective of digital
government. To understand the intertwined forces that play a role in this transformation process, and their
dynamics, contributions from disparate fields and discourses on this topic should be contrasted and
compared.
At the same time, the Communication on “Building the data economy” (COM (2017) 9) looks at proven or
potential blockages to the free movement of data and presents options to remove unjustified and or
disproportionate data location restrictions in the EU. It also considers the barriers around access to, and
transfer of, non-personal machine-generated data and data liability, as well as issues related to the
portability of non-personal data, interoperability and standards. In particular, it calls for the fostering of
technical solution development for the reliable identification and exchange of data.
A further avenue for research, complementary to this context of the digital transformation of government, is
the use of Web Application Programming Interfaces (hereafter called “APIs”)). APIs can be seen as “safe
entry ports for new and innovative uses of data” held by companies and, potentially, public administrations.
An opportunity exists to understand the current context and uptake of this technology early in the innovation
cycle of e-government in Europe, including for, but not limited to, geospatial data. 

In this context, Gartner was mandated by the Joint Research Centre (JRC) to conduct a study to provide an
analysis of the Web APIs as enablers for the digital transformation of governments. While digital
transformation of government is much wider than the technologies which can potentially support it, an
analysis of the role of APIs in the public sector is highly relevant to illustrate how technology can enable
transformation of government.
This study explores APIs, and their role in the EU public sector. It is divided up into the following sections:
• Introduction.
• Methodology: A description of the way in which this study has been conducted using different
research designs, including desk based research, consultation of API experts and multiple-case
study investigation. Also, using diverse research methods, including API reference and specific
document review, and interviews.
• Use of APIs in the Public Sector: Explores typical uses of APIs, such as to enable ecosystems,
before looking at some specific examples of API use.
• Relationship with location data: Explores how adding an additional layer of location data, with
other data (via APIs) can provide powerful use cases in many settings such as Smart Cities and
Gazetteers.
• Differences with the Private Sector: Explores any differences which may exist between the
public sector and private sector use of APIs.
• Future trends for API use: A view on how APIs may evolve, and their future uses in a world
where AI and Robotics are becoming prevalent.
• Case Study Insights – Presentation of the findings from case study interviews and associated
desk based research, plus analysis.
• Conclusions and Recommendations.
• Conclusion.

\clearpage

%********************************** %First Section  **************************************
\section{Glossary} %Section - 1.1 

% Note: It may be necessary to compile the document several times to get a multi-page table to line up properly
\begin{longtable}[c]{|l|l|}
	\hline
	\multicolumn{1}{|c|}{\textbf{Term}} & \multicolumn{1}{c|}{\textbf{Definition}} \\ \hline
	\endfirsthead
	%
	\endhead
	%
	API & \begin{tabular}[c]{@{}l@{}}Application Programming Interface - It is a set of clearly\\ defined methods of communication between the service\\ and any other software or components.\end{tabular} \\ \hline
	API Ecosystem & \begin{tabular}[c]{@{}l@{}}The developers, and the users of the application constructs\\ they build through an API, either within a company or on\\ the Internet with business partners, customers, citizens etc.\end{tabular} \\ \hline
	API Economy & \begin{tabular}[c]{@{}l@{}}A set of business models and channels — based on secure\\ access of functionality and exchange of data to an ecosystem\\ of developers and the users of the app constructs they\\ build — through an API, either within a company or on the\\ Internet with business partners, customers, citizens etc.\end{tabular} \\ \hline
	API Versioning & \begin{tabular}[c]{@{}l@{}}The ability to change without rendering older versions of\\ the same API inoperable.\end{tabular} \\ \hline
	\begin{tabular}[c]{@{}l@{}}API\\ Standardisation\end{tabular} & \begin{tabular}[c]{@{}l@{}}A uniform way for APIs to be expressed and consumed,\\ from COM and CORBA object brokers to web services\\ to today’s RESTful patterns.\end{tabular} \\ \hline
	\begin{tabular}[c]{@{}l@{}}API information\\ control\end{tabular} & \begin{tabular}[c]{@{}l@{}}A built-in means for enriching and handling the\\ information embodied by the API. This information\\ includes metadata, approaches to handling batches of\\ records, and hooks for middleware platforms, message\\ brokers, and service buses. It also defines how APIs\\ communicate, route, and manipulate the information\\ being exchanged.\end{tabular} \\ \hline
	API portal & \begin{tabular}[c]{@{}l@{}}A means for developers to discover, collaborate,\\ consume, and publish APIs. To support the overall\\ goal of self-service, these portals describe APIs in\\ a way that represents their functionality, context\\ (the business semantics of what they do, and how \\ they do it), non-functional requirements (scalability,\\ security, response times, volume limits, and resiliency\\ dimensions of the service), versioning, and metrics \\ tracking usage, feedback, and performance.\\ For organizations without mature master data or\\ architectural standards, the API portal can still offer\\ visibility into existing APIs and provide contact information\\ for individuals who can describe features, functions, and\\ technical details of services.\end{tabular} \\ \hline
	API gateway & \begin{tabular}[c]{@{}l@{}}A mechanism that allows consumers to become\\ authenticated and to “contract” with API specifications\\ and policies that are built into the API itself. Gateways\\ make it possible to decouple the “API proxy”—the node\\ by which consumers logically interact with the service—from\\ the underlying application for which the actual service is \\ being implemented. The gateway layer may offer the means\\ to load balance and throttle API usage.\end{tabular} \\ \hline
	API brokers & \begin{tabular}[c]{@{}l@{}}Enrichment, transformation, and validation services\\ to manipulate information coming to/from APIs, as\\ well as tools to embody business rule engines, workflow,\\ and business process orchestration on top of underlying APIs.\end{tabular} \\ \hline
	\begin{tabular}[c]{@{}l@{}}API\\ management\\ and monitoring\end{tabular} & \begin{tabular}[c]{@{}l@{}}A centralized and managed control level that provides\\ monitoring, service level management, SDLC process\\ integration, and role-based access management across\\ all three layers above. It includes the ability to instrument\\ and measure API usage, and even capabilities to price and\\ bill charge-back based on API consumption—to internal,\\ or potentially external, parties.\end{tabular} \\ \hline
	RESTful API & \begin{tabular}[c]{@{}l@{}}REST stands for “representational state transfer.” APIs\\ built according to REST architectural standards are\\ stateless and offer a simpler alternative to some SOAP\\ standards. For example, REST enables plain-text exchanges\\ of data assets instead of using complex WSDL protocols.\\ It also makes it possible to inherit security policies\\ from an underlying transport mechanism. At a high level,\\ these and other simplified approaches can deliver better\\ performance and faster paths to develop, deploy, and\\ triage.\end{tabular} \\ \hline
	\caption{}
	\label{glossary}\\
\end{longtable}

%********************************** %Second Section  *************************************
\section{API Overview} %Section - 1.2


APIs have become a foundational technological component of modern digital architectures, impacting every
sector of the global economy. In the public sector specifically, APIs are a key enabler of the accelerated
evolution of government and its agencies from analogue (manual, paper) operations, to digital.
The purpose of our study is to create a report that will support member states with the adoption of APIs
when in pursuit of their digital transformation. In order to explore this purpose, our investigation has
incorporated (but is not limited to) the following topics:
• The current use of APIs in the EU public sector.
• Differences between API use in the public sector and the private sector.
• The future trends for APIs.
• The relationship with location data.
• Insights from eight case study interviews, using real world examples to explore different
dimensions of the API landscape, including API Strategy, API Ecosystems, and specific APIs.
Our approach is covered in detail in the Methodology section later in this report. In summary, we have used
a combination of desk based research and interviews to gather information for analysis. For the interviews,
we used a questionnaire to collect information from representatives of a set of successful but diverse API
based case studies from a range of EU countries and sectors.
API Overview
API interaction occurs when one application would like to:
• Access or query the data held by another application
• Send data to that application
• Update data held in that application
• Request a service from another application
Types of APIs
Crucially, the use of APIs therefore simplifies, and standardises the interface reducing complexity and cost
of deployment over that of custom built interfaces. APIs represent an architectural approach that revolves
around providing programmable interfaces to different applications. It is technology agnostic, and creates a
flexible, loosely coupled architecture that allows a solution to be made up of components that can more
easily be switched in and out. The API approach is also a key enabler for application developers to build
apps that rapidly adapt to end user needs.
In the public sector, APIs enable important functionality and information held in one agency’s system or
department to be readily available to another without significant and expensive development effort. As well
as cross-departmental (agency) access to functionality and information (or even cross-border with a
different country’s administration) APIs also provide the ability to share information and functionality more
widely, i.e. to developers and ultimately to citizens for consumption through web or mobile based
applications.
Although there are many different types of API (see Appendix I), this study is most concerned with Web
APIs. Web services expose these APIs as endpoints that any internet-enabled language or software can
access, in exactly the same way browsers access websites and services7
. Web APIs deliver requests to the
service provider, and then deliver the response back to the requestor, i.e. they are an interface for web
applications, or applications that need to connect to each other via the Internet to communicate8
.
Web APIs themselves can be broken down further based on the type of data format that they harness, for
example, well known types are Simple Object Access Protocol (SOAP), Remote Procedure Call (RPC)
based APIs, and the Representational State Transfer (REST) architectural style. GraphQL – is a data query
language growing in popularity and has been adopted by leading social media outlets such as Facebook
and Pinterest9 as a type of API. While typical REST APIs require loading from multiple URLs, GraphQL
APIs get all the data an app developer needs in a single request enhancing speed of response even on
slow mobile network connections10.
Whilst the more traditional APIs are used as integration points within systems hidden from view, Web APIs
are often publicly available and can be ‘advertised’ via API Directory sites online. Tens of thousands11 are
available for developers to deliver consumable information to end users to do everything, from checking
traffic and weather, to updating a social media status, or even to make payments.
In the geospatial domain, besides existing private companies famous API proposals (e.g. Google Map), the
Open Geospatial Consortium (OGC) has created standards to support the exchange of geospatial
information12. They describe their Web services API standards as an agreed specification of rules and
guidelines about how to implement software interfaces and data encodings13. Geospatial software vendors,
developers and users collaborate in the OGC’s consensus process to develop and agree on standards that
enable information systems to exchange geospatial information and instructions for geoprocessing. OGC
standards are open standards. The OGC interface standards are also available in the REST style, and
cover a number of aspects:
• Visualisation standards e.g. Web Map Service (WMS).
• Data Access Standards e.g. Web Feature Service (WFS), SensorThings API.
• Processing Standards e.g. Web Processing Service (WPS).
• Metadata and Catalogue Service Standards e.g. Catalog Service for the Web (CWS).
• The informatics contract between the client code which manipulates normalized data structures of
geographic information based on the published API and the library code, e.g. the GeoAPI
Implementation Standard.
The standards above are part of the few globally agreed specifications adopted by the Technical Committee
211 of the International Organization for Standardization (ISO). ISO is also known to be working on
standards in other sectors, notably in Financial Services with ISO 2002214: however, because of they are
work in progress, details about them are still limited.
Whilst standards of this formal and specific nature are used in the EU, there is clear evidence that the need
for harmonising APIs lifecycle has been recognised. For example, the UK Government Digital Service
recognised that departments were developing APIs using different tools, platforms and approaches15, and
have set about working with industry to create a set of common principles for API design. The output has
been a set of guidelines on how developers working in any UK public sector organisation should build
APIs16 to ensure consistency, and success. Although they are titled as a ‘standard’, they are generic, and
not exact or specific in the way that an ISO or OGC standard is. Nevertheless, given the fact that
government is increasingly using APIs to automate processes and provide citizens with access to new
services it is hoped this approach will make integration simpler and faster.